\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{October 2}
\newcommand{\assignmentassignedate}{September 25}
\newcommand{\assignmentnumber}{Four}

\newcommand{\labyear}{2018}
\newcommand{\labday}{Tuesday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

% NOTE: The source code is simplified now that we use JavaFX

\newcommand{\mainprogram}{\lstinline{DisplayGraphicalScene}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labfour/DisplayGraphicalScene.java}}
\newcommand{\secondprogram}{\lstinline{DisplayGraphicalScene}}
\newcommand{\secondprogramsource}{\lstinline{src/main/java/labfour/DisplayGraphicalScene.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labfour/DisplayGraphicalScene.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labfour/DisplayGraphicalScene.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

% Enable margin notes to catch student attention

\usepackage{marginnote}
\reversemarginpar
\renewcommand*{\raggedrightmarginnote}{\centering}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}
\usepackage{fontawesome}

% Commands for margin notes, for optional use

\newcommand{\book}{\faBook}
\newcommand{\group}{\faUsers}

% Commands for making key margin notes, fully integrated

\newcommand{\caution}[1]{\null\hfill\LARGE{\faWarning{}}\newline\scriptsize{\em{#1}}}
\newcommand{\discuss}[1]{\null\hfill\LARGE{\faCommentO{}}\newline\scriptsize{\em{#1}}}
\newcommand{\resource}[1]{\null\hfill\LARGE{\faLink{}}\newline\scriptsize{\em{#1}}}
\newcommand{\think}[1]{\null\hfill\LARGE{\faCogs{}}\newline\scriptsize{\em{#1}}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 100}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 100\\Introduction to Computer Science I\\
      Fall 2018\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to practice using GitHub to access the files for a laboratory
assignment. Additionally, to master the use of the Ubuntu operating system and
software development programs such as a ``terminal window'' and a ``text
editor''. You will also continue to practice using Slack to support
communication with the teaching assistants and the course instructor. Next, you
will learn how to implement a Java program that creates graphical objects and
displays them on the user's screen. Finally, you will continue to explore the
use of methods, variables, and data types in Java.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory
  feature specialized software for completing this course's laboratory and
  practical assignments. If it is necessary for you to work on a different
  machine, be sure to regularly transfer your work to a laboratory machine so
  that you can check its correctness. If you cannot use a laboratory computer
  and you need help with the configuration of your own laptop, then please
  carefully explain its setup to a teaching assistant or the course instructor
  when you are asking questions.

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a teaching
  assistant or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a teaching assistant or the course instructor.

\item {\bf Store your files in GitHub}. As in the past laboratory assignments,
  you will be responsible for storing all of your files (e.g., Java source code
  and Markdown-based writing) in a Git repository using GitHub Classroom. Please
  verify that you have saved your source code in your Git repository by using
  \command{git status} to ensure that everything is updated. You can see if your
  assignment submission meets the established correctness requirements by using
  the provided checking tools on your local computer and in checking the commits
  in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Back up your files regularly}. All of your files are regularly
  backed-up to the servers in the Department of Computer Science and, if you
  commit your files regularly, stored on GitHub. However, you may want to use a
  flash drive, Google Drive, or your favorite backup method to keep an extra
  copy of your files on reserve. In the event of any type of system failure, you
  are responsible for ensuring that you have access to a recent backup copy of
  all your files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the teaching assistants and
  tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the tutors, teaching assistants, and
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should also
review Chapter 1 through Chapter 3 in the course textbook, paying close
attention to Sections 3.9 through 3.11 and Figures 3.11 and 3.14, and Listing
3.9. See the instructor if you have questions these readings.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into your web browser.
Now, you should accept the laboratory assignment and see that GitHub Classroom
created a new GitHub repository for you to access the assignment's starting
materials and to store the completed version of your assignment. Specifically,
to access your new GitHub repository for this assignment, please click the green
``Accept'' button and then click the link that is prefaced with the label ``Your
assignment has been created here''. If you accepted the assignment and correctly
followed these steps, you should have created a GitHub repository with a name
like
``Allegheny-Computer-Science-100-Fall-2018/computer-science-100-fall-2018-lab-4-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into the right directory you
should now type \command{cd cs100F2018}. Next, you can type the command
\command{ls} and see that there are some files or directories inside of this
directory. By typing \command{git clone} in your terminal and then pasting in
the string that you copied from the GitHub site you will download all of the
code for this assignment. For instance, if the course instructor ran the
\command{git clone} command in the terminal, it would look like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-100-F2018/computer-science-100-fall-2018-lab-4-gkapfham.git
\end{lstlisting}

\marginnote{\discuss{Discuss graphics code}}[.225in] After this command
finishes, you can use \command{cd} to change into the new directory. If you want
to \step{go back} one directory from your current location, then you can type
the command \command{cd ..}. Please continue to use the \command{cd} and
\command{ls} commands to explore the files that you automatically downloaded
from GitHub. What files and directories do you see? What do you think is their
purpose? Spend some time exploring, sharing your discoveries with a neighbor and
a \mbox{teaching assistant}. Now, try to answer the following questions, based
on Chapter 3, about how to create graphics with JavaFX.

\begin{enumerate}

  \setlength{\itemsep}{0pt}

  \item Can you draw a picture of the coordinate system used by JavaFX?

  \item How do you locate a point on the coordinate system used by JavaFX?

  \item How do you declare and display a line that runs from the point $(10,
    20)$ to the point $(300, 80)$?

  \item How do you declare and display a $200 \times 70$ rectangle that has
    a upper-left corner at $(30, 50)$?

  \item In JavaFX, what are the similarities and differences between a circle
    and an ellipse?

  \item How do you declare and display a circle with radius $40$ and a center
    point at $(100, 150)$?

  \item Using the JavaFX framework, how do you declare and display an ellipse
    with a center point at $(100, 50)$, a horizontal radius of $80$, and
    vertical radius of $30$?

  \item How do you group together (and place on a scene) different JavaFX shapes
    (e.g., a \program{Circle})?

  \item How do you define the color of a single graphical object like a
    \program{Circle} or an \program{Ellipse}?

  \item How do you set the title of your graphical display and then cause it to
    appear on the screen?

\end{enumerate}

\section*{Displaying Sophisticated Graphical Output}

In order view the source code of a Java program or a Markdown-based writing
assignment, you need a text editor. There are many different text editors on
your workstation and you should feel free to explore these on your own. Since it
is a powerful text editor known for helping computer scientists ``edit text at
the speed of thought'', in this class some students may opt to use the text
editor called \command{gvim}. Today, you will view and improve another Java
program in {\tt gvim}! To load the source code of today's Java program into
\program{gvim}, you should type \command{gvim
src/main/java/labfour/DisplayGraphicalScene.java} in your terminal window. As an
alternative to using \command{gvim}, students may also take similar steps to
view and edit the Java source code with the Atom text editor.

This assignment invites you to implement desktop wallpaper art similar to the
examples that are found on the web site called \url{http://simpledesktops.com/}.
Please see the Java source code comments for more details about the program's
purpose! The remainder of this section summarizes the goals for \mbox{this
program}. This assignment asks you to write a program, \mainprogramsource{},
that will display a canvas containing a graphical scene. For instance, it must
meet these requirements (see the README in your GitHub repository for more
details):

\vspace*{-.1in}
\begin{enumerate}
  \setlength{\itemsep}{0pt}

  \item Declare and use the constant variable called \command{CANVAS_HEIGHT}.

  \item Declare and use the constant variable called \command{CANVAS_WIDTH}.

  \item Declare and display at least five instances of the \program{Circle} class.

  \item Declare and display at least three instance of the \program{Line} class.

  \item Declare and display at least three instances of the \program{Rectangle}
    class.

  \item Declare and display at least two instances of the \program{Line} class.

  \item Display a graphics window containing the label ``\program{A
    Sophisticated Graphical Scene}''.

  \item Display a non-random, fully organized graphical scene suitable for a
    classroom setting.

\end{enumerate}

% Whenever you are finished typing text Java code in \program{gvim}, press the ESC
% key located in the upper left corner of the keyboard. This should remove the
% word \command{--INSERT--} from the bottom of the screen and take you out of
% insert mode. Use the \option{File/Save} menu to save your program.
% Alternatively, if you would like to use the keyboard to save your file, you can
% press \command{:w} when you are not in insert mode. Leaving the {\tt gvim}
% window open, go back to your terminal window.

If you want to \step{build} your program you can type the command \gradlebuild{}
in your terminal, thereby causing the Java compiler to check your program for
errors and get it ready to run. If you get any error messages, go back into your
text editor and try to figure out what you mis-typed and fix it. Once you have
solved the problem, make a note of the error and the solution for resolving it.
Re-save your program and then build it again by re-running the \gradlebuild{}.
If you cannot build \mainprogram{} correctly, then please talk with the course
instructor. Remember, you can produce any type of graphical display as long as
it meets the assignment's requirements.

\marginnote{\think{Reflect on challenges}}[.5in] When all of the errors are
eliminated, you can run your program by typing \gradlerun{} in the terminal
window---this is the ``execute'' step that will run your program and produce the
designated output. You should see no terminal output and instead notice the
graphical display. Make sure that all aspects of the graphics (e.g., their
position, size, and color) display correctly. If they do not, then repair the
program and re-build and re-run it. Once the
program runs and produces the intended output, please reflect on this process.
What step did you find to be the most challenging? Why? You should write your
reflections in a file, called \reflection{}, that uses the Markdown writing
language. To complete this aspect of the assignment, you should write one
high-quality paragraph that reports on your experiences with the various
commands and Java code segments.

\section*{Checking the Correctness of Your Program and Writing}

As verified by Checkstyle, the code for the \mainprogramsource{} and the \secondprogramsource{} files must adhere to all
of the requirements in the Google Java Style Guide available at
\url{https://google.github.io/styleguide/javaguide.html}. The Markdown file that contains your reflection must adhere to
the standards described in the Markdown Syntax Guide \url{https://guides.github.com/features/mastering-markdown/}.
Finally, your \reflection{} file should adhere to the Markdown standards established by the \step{Markdown linting} tool
available at \url{https://github.com/markdownlint/markdownlint/} and the writing standards set by the \step{prose
linting} tool from \url{http://proselint.com/}. Instead of requiring you to manually check that your deliverables adhere
to these industry-accepted standards, the GatorGrader tool that you will use in this laboratory assignment makes it easy
for you to automatically check if your submission meets these well-established standards for correctness.

Since this is not your first laboratory assignment, you will notice that the
provided source code does not contain all of the required comments in Java
source code file. This means that you will have to inspect the source code from
previous laboratory and practical assignments to review how to add comments to
the \mainprogramsource{}. Moreover, the provided source code is missing many of
the lines that are needed to pass the GatorGrader checks. Please review the
requirements for this Java source code file, as outlined in the previous
section. You should also study the code in this file to learn more about what
you need to add to it. Please don't forget to remove a \program{TODO} marker
from the code after you have fully resolve it!

\marginnote{\caution{Write useful commits}}[.75in] To get started with the use
of GatorGrader, type the command \gatorgraderstart{} in your terminal window. If
your laboratory work does not meet all of the assignment's requirements, then
you will see a summary of the failing checks along with a statement giving the
percentage of checks that are currently passing. If you do have mistakes in your
assignment, then you will need to review GatorGrader's output, find the mistake,
and try to fix it. Once your program is building correctly, fulfilling at least
some of the assignment's requirements, you should transfer your files to GitHub
using the \gitcommit{} and \gitpush{} commands. For example, if you want to
signal that the \mainprogramsource{} file has been changed and is ready for
transfer to GitHub you would first type \gitcommitmainprogram{} in your
terminal, followed by typing \gitpush{} and then checking to see that the
transfer to your GitHub repository is successful.

% To get started with the use of GatorGrader, type the command \gatorgraderstart{}
% in your terminal window. Once this step completes you can type
% \gatorgradercheck{}. If your work does not meet all of the assignment's
% requirements, then you will see the following output in your terminal:
% \command{Overall, are there any mistakes in the assignment? Yes}. If you do have
% mistakes in your assignment, then you will need to review GatorGrader's output,
% find the mistake, and try to fix it. Once your program is building correctly,
% fulfilling at least some of the assignment's requirements, you should transfer
% your files to GitHub using the \gitcommit{} and \gitpush{} commands. For
% example, if you want to signal that the \mainprogramsource{} file has been
% changed and is ready for transfer to GitHub you would first type
% \gitcommitmainprogram{} in your terminal, followed by typing \gitpush{}, and
% then checking to see that the transfer to GitHub is successful. Remember, to
% correctly complete this assignment you need to commit all code and writing files
% to GitHub. If you notice that the network communication with GitHub did not
% work, then please try to determine why, asking a teaching assistant or the
% instructor for assistance.

After the course instructor enables \step{continuous integration} with a system
called Travis CI, when you use the \gitpush{} command to transfer your source
code to your GitHub repository, Travis CI will initialize a \step{build} of your
assignment, checking to see if it meets all of the requirements. If both your
source code and writing meet all of the established requirements, then you will
see a green \checkmark{} in the listing of commits in GitHub after awhile. If
your submission does not meet the requirements, a red \naughtmark{} will appear
instead. The instructor will reduce a student's grade for this assignment if the
red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit
in your GitHub repository, then you satisfied all of the main checks, thereby
allowing the course instructor to evaluate other aspects of your source code and
writing, as further described in the \step{Evaluation} section of this
assignment sheet. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment. In conclusion, here are some
points to remember for creating programs that display graphical output:

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item You should think carefully about how your graphics can be displayed using
  variables.

\item As in past assignments, your program only needs to have one {\tt main}
  method in one file.

\item See Figure 3.12 for an illustrative example of the hierarchy for the
  display of graphical objects.

\item See Figure 3.14 for a listing of some common \command{Color} variables for
  use in your program.

\item If necessary, follow the book's instructions to use the RGB encoding to
  create your own \program{Color}.

\item Your program must use the \program{setTitle} method to give the required
  title to the graphic display.

\item Your final program should not contain any \program{TODO} markers as these
  are only for unfinished tasks.

\item Don't forget to review the assignment sheets from the previous laboratory
  and practical assignments as they contain insights that will support your
  completion of this assignment.

\end{enumerate}

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in \reflection{}, a one-paragraph reflection on the commands that
  you typed in \command{gvim} and the terminal window. This Markdown-based
  document should explain the input, output, and behavior of each command and
  the challenges that you confronted when using it. For every challenge that you
  encountered, please explain your solution for it.

\item A complete and correct version of \mainprogramsource{} that both meets all
  of the established requirements and produces the desired graphical output.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the requirements.
Other factors will also influence your final grade on the assignment. In
addition to studying the efficiency and effectiveness of your Java source code,
the instructor will also evaluate the accuracy of both your technical writing
and the comments in your source code. If your submission receives a red
\naughtmark{}, the instructor will reduce your grade for the assignment while
still considering the regularity with which you committed to your GitHub
repository and the overall quality of your partially completed work. Please see
the instructor if you have questions about the evaluation of this laboratory
assignment.

% \section*{Adhering to the Honor Code}

% In adherence to the Honor Code, students should complete this assignment on an
% individual basis. While it is appropriate for students in this class to have
% high-level conversations about the assignment, it is necessary to distinguish
% carefully between the student who discusses the principles underlying a
% problem with others and the student who produces assignments that are
% identical to, or merely variations on, someone else's work. Deliverables
% (e.g., Java source code or Markdown-based technical writing) that are nearly
% identical to the work of others will be taken as evidence of violating the
% \mbox{Honor Code}. Please see the course instructor if you have questions
% about this policy.

\end{document}
